---
title: "(Q3) Are traits coordinated along the “fast to slow” continuum and to what extent does this correlate with growth rates among species?"
author: "Main code: Leland Werden, Modified by: ECM"
date: "2021"
output: 
 prettydoc::html_pretty:
    highlight: pygments
    theme: cayman   
    toc: yes
    number_sections: False
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.align = 'center',
                      fig.width = 16,fig.height = 10)
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
require(vegan)
require(knitr)
require(ggpubr)
library(ggeffects)
library(emmeans)
library(MuMIn)
library(performance)
library(DHARMa)
library(car)
library(janitor)
library(factoextra)
library(nlme)
```


# Load and Clean Data

```{r}
final_trait_dat <- read.table("~/Documents/projects/shade_house_experiment/data/data_rda_leland.csv", 
                              header=TRUE, sep=",", strip.white=TRUE)
```


```{r}

# Select variables that are going to be used in the model
all_trait <-  
    final_trait_dat %>%
  
    # Select columns needed for regression
    dplyr::select(spcode, treatment, nfixer, init_height, 
                  
                  # Performance measures
                  totalbiom, above_biom,below_biom, agr, 
                  
                  # physiology traits
                  amax, gs, wue,
                  
                  # Leaf traits
                  Narea_g_m2,Nmass_mg_g,d13c, sla_cm2_g) %>% 
    
    # Convert to factors type columns
    mutate(spcode = factor(spcode) , 
           treatment = factor(treatment) , 
           nfixer = factor(nfixer) )



#Generate a copy of the original data
all_trait_scaled <- all_trait

# Scale and center traits
all_trait_scaled[,c(9:15)] <- scale(all_trait_scaled[,c(9:15)],center = TRUE, scale = TRUE)
```

# PCA 

```{r}
pca_final_traits <- princomp(all_trait_scaled[,c(9:15)])
```

```{r}
fviz_screeplot(pca_final_traits, addlabels = TRUE, ylim = c(0, 50))
```

```{r}
#Extract loadings
pca_final_traits$loadings 
```


## Biplots

### PC1 vrs PC2

```{r}

fviz_pca_biplot(pca_final_traits,axes = c(1, 2),
                
                # Add group 
                habillage=all_trait_scaled$nfixer, 
                
                # Set labels
                label ="var",
                
                # Set arrow color
                col.var = "black",
                
                # Repel overlapping labels
                repel = TRUE) +
  theme_minimal() + scale_color_brewer(palette="Set1")
```

### PC1 vrs PC3

```{r}
fviz_pca_biplot(pca_final_traits,axes = c(1, 3),
                
                # Add group 
                habillage=all_trait_scaled$nfixer, 
                
                # Set labels
                label ="var",
                
                # Set arrow color
                col.var = "black",
                
                # Repel overlapping labels
                repel = TRUE) +
  theme_minimal() + scale_color_brewer(palette="Set1")
```

### PC2 vrs PC3
```{r}
fviz_pca_biplot(pca_final_traits,axes = c(2, 3),
                
                # Add group 
                habillage=all_trait_scaled$nfixer, 
                
                # Set labels
                label ="var",
                
                # Set arrow color
                col.var = "black",
                
                # Repel overlapping labels
                repel = TRUE) +
  theme_minimal() + scale_color_brewer(palette="Set1")
```

## Biplots interpretation
[PCA interpretation](https://wilkelab.org/SDS375/slides/dimension-reduction-1.html#18)

+ The first 3 PCs explains 83% (47.9,19.4,15.8) of the variability in the data
+ PC1 separates fixers from non-fixers
+ All traits contribute positively to PC1 

+ PC1 represents the overall high photosynthetic capacity 
+ PC2 represents the difference between stomatal conductance and SLA
+ PC3 represents the difference between d13c and SLA

# Regression model with PCs as independent variables

```{r}
all_trait_scaled_for_reg <- all_trait_scaled
```


```{r}
# Pull out PC axes and add to trait/growth data frame
pca_scores <- scores(pca_final_traits)
```

```{r}
#Extract PCs
all_trait_scaled_for_reg$PC1 <- pca_scores[,1]
all_trait_scaled_for_reg$PC2 <- pca_scores[,2]
all_trait_scaled_for_reg$PC3 <- pca_scores[,3]
```


# Model formula used 

$$response\sim treatment*fixer\ + initial\ height\ + random(1|specie)$$

# Perform regression models with AGR as response variable

```{r}
# add t-ratio
emm_options(contrast = list(infer = c(TRUE, TRUE)))
```

## Do PCs predict AGR? varIdent Two-Way interaction model

```{r}
# fit treatment/nfixer interactions separately
varident_pca_mod_agr_two_way_comp <- lme(agr ~ treatment + nfixer+ PC1 + PC2 + PC3 +

                            #Two way interactions
                            #Treatments
                            treatment * PC1 +
                            treatment * PC2 +
                            treatment * PC3 +

                            #NFixer
                            nfixer * PC1 +
                            nfixer * PC2 +
                            nfixer * PC3 +

                            #Control over different plant heights
                            init_height,

                            #Random part
                            random = ~ 1|spcode,
                            
                            #use ML to compare models
                            method = "ML",
                            
                            #varIdent
                            weights = varIdent(form = ~ 1 | treatment*nfixer),

                        data = all_trait_scaled_for_reg)
```


## Do PCs predict AGR? Three-Way interaction model

```{r}
# fit three-way interaction
varident_pca_mod_agr_three_way_comp <- lme(agr ~ treatment * nfixer* PC1 +
                                            treatment * nfixer* PC2 +
                                            treatment * nfixer* PC3 +

                                            #Control over different plant heights
                                            init_height,

                                            #Random part
                                            random = ~ 1|spcode,
                                      
                                            #use ML to compare models
                                            method = "ML",
                            
                                            #varIdent
                                            weights = varIdent(form = ~ 1 | treatment*nfixer),

                        data = all_trait_scaled_for_reg)
```

## Model comparision and selection

```{r}
# check two-way vs. three-way interaction models 
#(shows that two-way is way better, more negative AICc -- include in methods)
compare_performance(varident_pca_mod_agr_two_way_comp, varident_pca_mod_agr_three_way_comp,
                    rank = TRUE,metrics = "all")
```

```{r}
anova(varident_pca_mod_agr_two_way_comp,varident_pca_mod_agr_three_way_comp)
```

## Model selected

```{r}
varident_pca_mod_agr_two_way <- lme(agr ~ treatment + nfixer+ PC1 + PC2 + PC3 +

                            #Two way interactions
                            #Treatments
                            treatment * PC1 +
                            treatment * PC2 +
                            treatment * PC3 +

                            #NFixer
                            nfixer * PC1 +
                            nfixer * PC2 +
                            nfixer * PC3 +

                            #Control over different plant heights
                            init_height,

                            #Random part
                            random = ~ 1|spcode,
                            
                            # Changed to REML
                            method = "REML",
                            
                            #varIdent
                            weights = varIdent(form = ~ 1 | treatment*nfixer),

                        data = all_trait_scaled_for_reg)
```



## ANOVA table
```{r message=FALSE, warning=FALSE}
# ANOVA -- PC1 X Treatment effect significant (chi-sq = 29.25, P <0.001) 
anova.lme(varident_pca_mod_agr_two_way, type = "marginal", adjustSigma = F) %>% 
  kable()
```

## PCA trends

### PC1 trends contrasts

```{r}
# compare slopes of lines for treatment X PC1 iteraction
# summary(emtrends(model = fit_sat, ~agef, var="SATV"), infer=TRUE)
#emtrends(pca_mod_agr_1, pairwise ~ treatment|PC1, var = "PC1")

emtrends(varident_pca_mod_agr_two_way,pairwise ~ treatment, var = "PC1")$contrast  %>% 
  kable()
```

### PC1 trends plot

```{r}
# plot that shit
cbbPalette <- c("#F0E442", "#009E73", "#56B4E9", "#0072B2")

PC1_predict <- ggpredict(varident_pca_mod_agr_two_way, terms = c("PC1","treatment"))

(PC1_AGR_fig <- ggplot(PC1_predict, aes(x, predicted, colour = group)) + 
                      geom_line()+
                      ylab("AGR")+
                      xlab("PC1")+ 
                      theme_classic()+
                      scale_color_manual(values= cbbPalette,
                                        labels=c("No addition", "Plus nutrients addition", 
                                                  "Plus water addition", 
                                                 "Plus nutrients and water addition")
                                        ) +
                      geom_line(size=1))


```

### PC2 trends contrasts

```{r}
# compare slopes of lines for treatment X PC1 iteraction
emtrends(varident_pca_mod_agr_two_way, pairwise ~ treatment, var = "PC2")$contrast %>% 
  kable()
```

### PC2 trends plot

```{r}
cbbPalette <- c("#F0E442", "#009E73", "#56B4E9", "#0072B2")

PC2_predict <- ggpredict(varident_pca_mod_agr_two_way, terms = c("PC2", "treatment"))

(PC2_AGR_fig <- ggplot(PC2_predict, aes(x, predicted, colour = group)) + 
                      geom_line()+
                      ylab("AGR")+
                      xlab("PC2")+ 
                      theme_classic()+
                      scale_color_manual(values= cbbPalette,
                                        labels=c("No addition", "Plus nutrients addition", 
                                                  "Plus water addition", 
                                                 "Plus nutrients and water addition")
                                        )+
                      geom_line(size = 1))

```
`
# Perform regression models with Total Biomass as response variable

## Do PCs predict Total Biomass? Two-Way interaction

```{r}
# fit treatment/nfixer interactions separately
# varident_pca_mod_totalbiom_two_way <- lme(totalbiom ~ treatment + nfixer+ PC1 + PC2 + PC3 +
# 
#                             #Two way interactions
#                             #Treatments
#                             treatment * PC1 +
#                             treatment * PC2 +
#                             treatment * PC3 +
# 
#                             #NFixer
#                             nfixer * PC1 +
#                             nfixer * PC2 +
#                             nfixer * PC3 +
# 
#                             #Control over different plant heights
#                             init_height,
# 
#                             #Random part
#                             random = ~ 1|spcode,
#                             
#                             method = "REML",
#                             
#                             #varIdent
#                             weights = varIdent(form = ~ 1 | treatment*nfixer),
# 
#                         data = all_trait_scaled_for_reg)
# 

```


## Do PCs predict Total Biomass? Three-Way interaction

```{r}
# fit three-way interaction
# varident_pca_mod_totalbiom_three_way <- lme(totalbiom ~ treatment * nfixer* PC1 +
#                                             treatment * nfixer* PC2 +
#                                             treatment * nfixer* PC3 +
# 
#                                             #Control over different plant heights
#                                             init_height,
# 
#                                             #Random part
#                                             random = ~ 1|spcode,
#                                             
#                                             method = "REML",
#                             
#                                             #varIdent
#                                             weights = varIdent(form = ~ 1 | treatment*nfixer),
# 
#                         data = all_trait_scaled_for_reg)
```

## Model comparision and selection

```{r}
# check two-way vs. three-way interaction models 
# (shows that two-way is way better, more negative AICc -- include in methods)
#compare_performance(varident_pca_mod_totalbiom_two_way, varident_pca_mod_totalbiom_three_way,metrics = "all",rank = TRUE)
```

```{r}
#anova.lme(varident_pca_mod_totalbiom_two_way, varident_pca_mod_totalbiom_three_way)
```


## ANOVA table
```{r message=FALSE, warning=FALSE}
# ANOVA -- PC1 X Treatment effect significant (chi-sq = 29.25, P <0.001) 
# (varident_pca_mod_agr_two_way, ) %>% 
  
#anova.lme(varident_pca_mod_totalbiom_three_way, type = "marginal", adjustSigma = F) %>% 
#  kable()
```


# Perform regression models with Belowground biomass as response variable

## Do PCs predict Belowground biomass? varIdent Two-Way interaction model

```{r}
# fit treatment/nfixer interactions separately
# varident_pca_mod_below_biom_two_way <- lme(below_biom ~treatment + nfixer+ PC1 + PC2 + PC3 +
# 
#                             #Two way interactions
#                             #Treatments
#                             treatment * PC1 +
#                             treatment * PC2 +
#                             treatment * PC3 +
# 
#                             #NFixer
#                             nfixer * PC1 +
#                             nfixer * PC2 +
#                             nfixer * PC3 +
# 
#                             #Control over different plant heights
#                             init_height,
# 
#                             #Random part
#                             random = ~ 1|spcode,
#                             
#                             method = "REML",
#                             
#                             #varIdent
#                             weights = varIdent(form = ~ 1 | treatment*nfixer),
# 
#                         data = all_trait_scaled_for_reg)
```


## Do PCs predict Belowground biomass? Three-Way interaction model

```{r}
# fit three-way interaction
# varident_pca_mod_below_biom_three_way <- lme(below_biom ~ treatment * nfixer* PC1 +
#                                             treatment * nfixer* PC2 +
#                                             treatment * nfixer* PC3 +
# 
#                                             #Control over different plant heights
#                                             init_height,
# 
#                                             #Random part
#                                             random = ~ 1|spcode,
#                                       
#                                             method = "REML",
#                             
#                                             #varIdent
#                                             weights = varIdent(form = ~ 1 | treatment*nfixer),
# 
#                         data = all_trait_scaled_for_reg)
```


## Model comparision and selection

```{r}
# check two-way vs. three-way interaction models 
#(shows that two-way is way better, more negative AICc -- include in methods)
# compare_performance(varident_pca_mod_below_biom_two_way, 
#                     varident_pca_mod_below_biom_three_way,
#                     rank = TRUE,metrics = "all")
```

```{r}
#anova.lme(varident_pca_mod_below_biom_two_way,varident_pca_mod_below_biom_three_way)
```


## ANOVA table
```{r message=FALSE, warning=FALSE}
# ANOVA -- PC1 X Treatment effect significant (chi-sq = 29.25, P <0.001) 
#anova.lme(varident_pca_mod_below_biom_two_way, type = "marginal", adjustSigma = F) %>% 
#  kable()
```

# Perform regression models with Aboveground biomass as response variable

## Do PCs predict Aboveground biomass? varIdent Two-Way interaction model

```{r}
# fit treatment/nfixer interactions separately
# varident_pca_mod_above_biom_two_way <- lme(above_biom ~treatment + nfixer+ PC1 + PC2 + PC3 +
# 
#                             #Two way interactions
#                             treatment * nfixer +
#                             #Treatments
#                             treatment * PC1 +
#                             treatment * PC2 +
#                             treatment * PC3 +
# 
#                             #NFixer
#                             nfixer * PC1 +
#                             nfixer * PC2 +
#                             nfixer * PC3 +
# 
#                             #Control over different plant heights
#                             init_height,
# 
#                             #Random part
#                             random = ~ 1|spcode,
#                             
#                             method = "REML",
#                             
#                             #varIdent
#                             weights = varIdent(form = ~ 1 | treatment*nfixer),
# 
#                         data = all_trait_scaled_for_reg)
```


## Do PCs predict Belowground biomass? Three-Way interaction model

```{r}
# fit three-way interaction
# varident_pca_mod_above_biom_three_way <- lme(above_biom ~ treatment * nfixer* PC1 +
#                                             treatment * nfixer* PC2 +
#                                             treatment * nfixer* PC3 +
# 
#                                             #Control over different plant heights
#                                             init_height,
# 
#                                             #Random part
#                                             random = ~ 1|spcode,
#                                       
#                                             method = "REML",
#                             
#                                             #varIdent
#                                             weights = varIdent(form = ~ 1 | treatment*nfixer),
# 
#                         data = all_trait_scaled_for_reg)
```


## Model comparision and selection

```{r}
# check two-way vs. three-way interaction models 
#(shows that two-way is way better, more negative AICc -- include in methods)
# compare_performance(varident_pca_mod_above_biom_two_way, 
#                     varident_pca_mod_above_biom_three_way,
#                     rank = TRUE,metrics = "all")
```

```{r}
#anova.lme(varident_pca_mod_below_biom_two_way,varident_pca_mod_below_biom_three_way)
```

## ANOVA table
```{r message=FALSE, warning=FALSE}
# ANOVA -- PC1 X Treatment effect significant (chi-sq = 29.25, P <0.001) 
# anova.lme(varident_pca_mod_above_biom_two_way, type = "marginal", adjustSigma = F) %>% 
#   kable()
```


