---
title: "(Q3) Are traits coordinated along the “fast to slow” continuum and to what extent does this correlate with growth rates among species?"
author: "Main code: Leland Werden, Modified by: Erick Calderon-Morales"
date: "2021"
output: 
 prettydoc::html_pretty:
    fig_width: 8
    fig_height: 7 
    highlight: pygments
    theme: cayman   
    toc: yes
    number_sections: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.align = 'center',
                      fig.width = 16,fig.height = 10)
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
require(vegan)
require(knitr)
require(ggpubr)
library(ggeffects)
library(emmeans)
library(lme4)
library(MuMIn)
library(performance)
library(DHARMa)
library(car)
library(janitor)
library(factoextra)
library(nlme)
```


# Load and Clean Data

```{r}
#Source maso menos plot
source("~/Documents/projects/shade_house_experiment/codes/masomenos_plot.R")
```


```{r}
final_trait_dat <- read.table("~/Documents/projects/shade_house_experiment/data/data_rda_leland.csv", header=TRUE, sep=",", strip.white=TRUE)
```


```{r}

# Select variables that are going to be used in the model
all_trait <-  
    final_trait_dat %>%
  
    # Select columns needed for regression
    dplyr::select(spcode, treatment, nfixer, init_height, 
                  
                  # Performance measures
                  totalbiom, above_biom,below_biom, agr, 
                  
                  # physiology traits
                  amax, gs, wue,
                  
                  # Leaf traits
                  Narea_g_m2,Nmass_mg_g,d13c, sla_cm2_g) %>% 
    
    # Convert to factors type columns
    mutate(spcode = factor(spcode) , 
           treatment = factor(treatment) , 
           nfixer = factor(nfixer) )



#Generate a copy of the original data
all_trait_scaled <- all_trait

# Scale and center traits
all_trait_scaled[,c(9:15)] <- scale(all_trait_scaled[,c(9:15)],center = TRUE, scale = TRUE)
```

# PCA 

```{r}
pca_final_traits <- princomp(all_trait_scaled[,c(9:15)])
```

```{r}
fviz_screeplot(pca_final_traits, addlabels = TRUE, ylim = c(0, 50))
```

```{r}
#Extract loadings
pca_final_traits$loadings 
```


## Biplots

### PC1 vrs PC2

```{r}

fviz_pca_biplot(pca_final_traits,axes = c(1, 2),
                
                # Add group 
                habillage=all_trait_scaled$nfixer, 
                
                # Set labels
                label ="var",
                
                # Set arrow color
                col.var = "black",
                
                # Repel overlapping labels
                repel = TRUE) +
  theme_minimal() + scale_color_brewer(palette="Set1")
```

### PC1 vrs PC3

```{r}

fviz_pca_biplot(pca_final_traits,axes = c(1, 3),
                
                # Add group 
                habillage=all_trait_scaled$spcode, 
                
                # Set labels
                label ="var",
                
                # Set arrow color
                col.var = "black",
                
                # Repel overlapping labels
                repel = TRUE) +
  theme_minimal() + scale_color_brewer(palette="Set1")
```

### PC2 vrs PC3
```{r}
fviz_pca_biplot(pca_final_traits,axes = c(2, 3),
                
                # Add group 
                habillage=all_trait_scaled$nfixer, 
                
                # Set labels
                label ="var",
                
                # Set arrow color
                col.var = "black",
                
                # Repel overlapping labels
                repel = TRUE) +
  theme_minimal() + scale_color_brewer(palette="Set1")
```

## Biplots interpretation
[PCA interpretation](https://wilkelab.org/SDS375/slides/dimension-reduction-1.html#18)

+ The first 3 PCs explains 83% (47.9,19.4,15.8) of the variability in the data
+ PC1 separates fixers from non-fixers
+ All traits contribute positively to PC1 

+ PC1 represents the overall high photosynthetic capacity 
+ PC2 represents the difference between stomatal conductance and SLA
+ PC3 represents the difference between d13c and SLA

# Regression model with PCs as independent variables

```{r}
all_trait_scaled_for_reg <- all_trait_scaled
```


```{r}
# Pull out PC axes and add to trait/growth data frame
pca_scores <- scores(pca_final_traits)
```

```{r}
#Extract PCs
all_trait_scaled_for_reg$PC1 <- pca_scores[,1]
all_trait_scaled_for_reg$PC2 <- pca_scores[,2]
all_trait_scaled_for_reg$PC3 <- pca_scores[,3]
```


# Model formula used 

$$response\sim treatment*fixer\ + initial\ height\ + random(1|specie)$$

# Perform regression models with PC1 as response variable

```{r}
# fit treatment/nfixer interactions separately
varident_pc1_mod <- lme(PC1 ~ treatment,
                          
                            #Random part
                            random = ~ 1|nfixer,
                            
                            #varIdent
                            weights = varIdent(form = ~ 1 | treatment*nfixer),

                        data = all_trait_scaled_for_reg)
```


## Model performance

```{r}
# check two-way vs. three-way interaction models 
#(shows that two-way is way better, more negative AICc -- include in methods)
model_performance(varident_pc1_mod,metrics = "all")

```

## ANOVA table
```{r message=FALSE, warning=FALSE}
# ANOVA -- PC1 X Treatment effect significant (chi-sq = 29.25, P <0.001) 
Anova(varident_pc1_mod, type = "III",test = "F") 
```



## PC1 contrasts

```{r message=FALSE, warning=FALSE}
data_pc1_plot <- 
  emmip(varident_pc1_mod, ~ treatment, CIs = TRUE)$data 

# Get by group comparitions
    as.data.frame(emmeans(varident_pc1_mod,
    					  pairwise ~ treatment,
    					  type = "response", 
    					  adjust ="tukey")$contrast) %>% 
  
    separate(contrast,c("group1","group2"), sep = " - " ) %>% 
    clean_names() %>% 
    mutate(group1 = factor(group1)) %>%
    mutate(group2 = factor(group2)) %>% 
    select(group1, group2,estimate,se,p_value)  %>%
    #filter(!group1 == "Harvestatthebegging") %>% 
    mutate(p_value = round(p_value, 20)) %>% 
    kable()
```

## PC1 plot

```{r}
(plot_totalbiom <- 
  masomenos_plot_no_lines(data = data_pc1_plot,xvar = xvar, yvar = yvar, 
                 tvar = tvar,n_treat = 4,
                 ucl = UCL, lcl = LCL) +
  theme(legend.position = "right") + ylab("PC1"))

```



# Perform regression models with PC2 as response variable

```{r}
# fit treatment/nfixer interactions separately
varident_pc2_mod <- lme(PC2 ~ treatment*nfixer,
                          
                            #Random part
                            random = ~ 1|spcode,
                            
                            #varIdent
                            weights = varIdent(form = ~ 1 | treatment*nfixer),

                        data = all_trait_scaled_for_reg)
```


## Model performance

```{r}
# check two-way vs. three-way interaction models 
#(shows that two-way is way better, more negative AICc -- include in methods)
model_performance(varident_pc2_mod,metrics = "all")
```

## ANOVA table
```{r message=FALSE, warning=FALSE}
# ANOVA -- PC1 X Treatment effect significant (chi-sq = 29.25, P <0.001) 
Anova(varident_pc2_mod, type = "III",test = "F") 
```

## PC2 contrasts

```{r message=FALSE, warning=FALSE}
data_pc2_plot <- 
  emmip(varident_pc1_mod, ~ treatment, CIs = TRUE)$data 

# Get by group comparitions
    as.data.frame(emmeans(varident_pc2_mod,
    					  pairwise ~ treatment,
    					  type = "response", 
    					  adjust ="tukey")$contrast) %>% 
  
    separate(contrast,c("group1","group2"), sep = " - " ) %>% 
    clean_names() %>% 
    mutate(group1 = factor(group1)) %>%
    mutate(group2 = factor(group2)) %>% 
    select(group1, group2,estimate,se,p_value)  %>%
    #filter(!group1 == "Harvestatthebegging") %>% 
    mutate(p_value = round(p_value, 20)) %>% 
    kable()
```

## PC2 plot

```{r}
(plot_pc2 <- 
  masomenos_plot_no_lines(data = data_pc2_plot,xvar = xvar, yvar = yvar, 
                 tvar = tvar,n_treat = 4,
                 ucl = UCL, lcl = LCL) +
  theme(legend.position = "right") + ylab("PC2"))

```


