---
title: "(Q3) Are traits coordinated along the “fast to slow” continuum and to what extent does this correlate with growth rates among species? Checking assumptions for PCA regression"

author: "Main code Leland Werden, edited by ECM"
date: "2021"
output: 
 prettydoc::html_pretty:
    fig_width: 8
    fig_height: 7 
    highlight: pygments
    theme: cayman   
    toc: yes
    number_sections: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.align = 'center',
                      fig.width = 16,fig.height = 10)
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
require(vegan)
require(knitr)
require(ggpubr)
library(ggeffects)
library(emmeans)
library(lme4)
library(MuMIn)
library(performance)
library(DHARMa)
library(car)
library(gridExtra)
library(janitor)
library(nlme)
library(factoextra)
```


# Load and Clean Data

```{r}
final_trait_dat <- read.table("~/Documents/projects/shade_house_experiment/data/data_rda_leland.csv",
                              header=TRUE, sep=",", strip.white=TRUE)
```

```{r}

# Select variables that are going to be used in the model
all_traits <-  
    final_trait_dat %>%
    dplyr::select(spcode, treatment, nfixer, init_height, 
                  
                  #Performance measures
                  totalbiom, above_biom,rgr, agr, 
                  
                  #physiology traits
                  amax, gs, wue,
                  
                  #Leaf traits
                  Narea_g_m2,Nmass_mg_g,d13c, sla_cm2_g)



#Generate a copy of the original data
all_trait_scaled <- all_traits

# Scale and center data traits
all_trait_scaled[,c(9:15)] <- scale(all_trait_scaled[,c(9:15)], center = TRUE, 
                                    scale = TRUE)
colnames(all_trait_scaled)
```


# PCA 
```{r}
pca_traits <- princomp(all_trait_scaled[,c(9:15)])
```

```{r}
#Extract loadings
pca_traits$loadings 
```

```{r}
# Pull out PC axes and add to trait/growth data frame
pca_scores <- scores(pca_traits)
```


# Data for PCA regression 
```{r}
all_trait_scaled_for_reg <- all_trait_scaled %>%
    
    #Transform to factor 
    mutate(
        nfixer = factor(nfixer),
        treatment = factor(treatment),
        spcode = factor(spcode)
    )


```

## Pull out PC axes and add to trait/growth data frame
```{r}
#Extract PCs
all_trait_scaled_for_reg$PC1 <- pca_scores[,1]
all_trait_scaled_for_reg$PC2 <- pca_scores[,2]
all_trait_scaled_for_reg$PC3 <- pca_scores[,3]
```

# Model

## Do PCs predict AGR? Two-Way interaction model: varIdent implemented
$$1)\ response\sim treatment*PC +\ fixer*PC\ + initial\ height\ + random(1|specie)$$


```{r message=FALSE, warning=FALSE}
# fit treatment/nfixer interactions separately
pca_mod_agr_1 <- lmer(agr ~ treatment + nfixer+ PC1 + PC2 + PC3 +
                            
                            #Control over different plant heights
                            init_height +  
                            
                            #Interactions
                            #Treament
                            treatment * PC1 +
                            treatment * PC2 +
                            treatment * PC3 +
   
                            #NFixer 
                            nfixer * PC1 +
                            nfixer * PC2 +
                            nfixer * PC3 +
                        
                            #Random part
                            (1|spcode), 
                        data = all_trait_scaled_for_reg)
```



```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted 
# as “the proportion of the variance explained by the grouping structure in the
# population”
model_performance(pca_mod_agr_1)
```


```{r message=FALSE, warning=FALSE, fig.align='center', fig.width=20, fig.height=10}

check_model(pca_mod_agr_1, panel = T) 
```


```{r}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_two_way_inter <- simulateResiduals(fittedModel = pca_mod_agr_1)
plot(simulation_output_two_way_inter) 
```


```{r}
check_heteroscedasticity(pca_mod_agr_1)
```


```{r}
plot(all_trait_scaled_for_reg$nfixer, resid(pca_mod_agr_1), xlab = "NFixer",
     ylab = "Residuals")
```

```{r}
plot(all_trait_scaled_for_reg$treatment, resid(pca_mod_agr_1), xlab = "Treatment",
	 ylab = "Residuals")
```


### First model PCA regression Two-way interaction

```{r}
pca_mod_agr_two_way <- lme(agr ~ treatment + nfixer+ PC1 + PC2 + PC3 +

                            #Two way interactions
                            #Treatments
                            treatment * PC1 +
                            treatment * PC2 +
                            treatment * PC3 +

                            #NFixer
                            nfixer * PC1 +
                            nfixer * PC2 +
                            nfixer * PC3 +

                            #Control over different plant heights
                            init_height,

                            #Random part
                            random = ~ 1|spcode,

                        data = all_trait_scaled_for_reg)
```

### varIdent PCA regression Two-way interaction

```{r}

varident_pca_mod_agr_two_way <- lme(agr ~treatment + nfixer+ PC1 + PC2 + PC3 +

                            #Two way interactions
                            #Treatments
                            treatment * PC1 +
                            treatment * PC2 +
                            treatment * PC3 +

                            #NFixer
                            nfixer * PC1 +
                            nfixer * PC2 +
                            nfixer * PC3 +

                            #Control over different plant heights
                            init_height,

                            #Random part
                            random = ~ 1|spcode,
                            
                            #varIdent
                            weights = varIdent(form = ~ 1 | treatment*nfixer),

                        data = all_trait_scaled_for_reg)




```


```{r}
anova(pca_mod_agr_two_way,varident_pca_mod_agr_two_way)
```

```{r}
plot(varident_pca_mod_agr_two_way,which = c(1), 
     col = all_trait_scaled_for_reg$treatment,add.smooth = T, 
     caption = "Treatment")
```

```{r}
plot(varident_pca_mod_agr_two_way,which = c(1), 
     col = all_trait_scaled_for_reg$nfixer,add.smooth = T, caption = "nfixer")
```


```{r}
E_two_way <- resid(varident_pca_mod_agr_two_way,type = "normalized")
coplot(E_two_way ~ treatment | nfixer, data = all_trait_scaled_for_reg, 
	   ylab = "Normalised residuals" )
```

## Do PCs predict AGR? Three-Way interaction model: varIdent implemented

$$2)\ response\sim treatment*fixer*PC + initial\ height\ + random(1|specie)$$

```{r message=FALSE, warning=FALSE}
# fit three-way interaction
pca_mod_agr_2 <- lmer(agr ~ treatment * nfixer* PC1 +
                            treatment * nfixer* PC2 +
                            treatment * nfixer* PC3 +
                            
                            #Control over different plant heights
                            init_height +
                            
                            #Random part
                            (1|spcode),
                      
                          data = all_trait_scaled_for_reg)
```


```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted 
# as “the proportion of the variance explained by the grouping structure in the
# population”
model_performance(pca_mod_agr_2)
```


```{r}
check_model(pca_mod_agr_2, panel = T) 
```


```{r message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_three_way_inter <- simulateResiduals(fittedModel = pca_mod_agr_2)
plot(simulation_output_three_way_inter) 
```


```{r}
check_heteroscedasticity(pca_mod_agr_2)
```


```{r}
plot(all_trait_scaled_for_reg$nfixer, resid(pca_mod_agr_2), xlab = "NFixer",
     ylab = "Residuals")
```

```{r}
plot(all_trait_scaled_for_reg$treatment, resid(pca_mod_agr_2), xlab = "Treatment",
	 ylab = "Residuals")
```

### First model PCA regression Three-way interaction

```{r}

pca_mod_agr_three_way <- lme(agr ~  treatment * nfixer* PC1 +
                                    treatment * nfixer* PC2 +
                                    treatment * nfixer* PC3 +

                                    #Control over different plant heights
                                    init_height,

                                    #Random part
                                    random = ~ 1|spcode,

                        data = all_trait_scaled_for_reg)
```

### varIdent PCA regression Three-way interaction

```{r}

varident_pca_mod_agr_three_way <- lme(agr ~ treatment * nfixer* PC1 +
                                            treatment * nfixer* PC2 +
                                            treatment * nfixer* PC3 +

                                            #Control over different plant heights
                                            init_height,

                                            #Random part
                                            random = ~ 1|spcode,
                            
                                            #varIdent
                                            weights = varIdent(form = ~ 1 | treatment*nfixer),

                        data = all_trait_scaled_for_reg)
```


```{r}
anova(pca_mod_agr_three_way,varident_pca_mod_agr_three_way)
```


```{r}
plot(varident_pca_mod_agr_three_way,which = c(1), 
     col = all_trait_scaled_for_reg$treatment,add.smooth = T, 
     caption = "Treatment")
```

```{r}
plot(varident_pca_mod_agr_three_way,which = c(1), 
     col = all_trait_scaled_for_reg$nfixer,add.smooth = T, caption = "nfixer")
```


```{r}
E <- resid(varident_pca_mod_agr_three_way,type = "normalized")
coplot(E ~ treatment | nfixer, data = all_trait_scaled_for_reg, 
	   ylab = "Normalised residuals" )
```

















