---
title: "(Q1) What is the relative influence of water vs. nutrient availability on tropical dry forest seedling performance. Checking assumptions for biomass variables"
author: "Mas o menos lab"
date: "2021"
output: 
 prettydoc::html_pretty:
    fig_width: 20
    fig_height: 12 
    highlight: pygments
    theme: cayman   
    toc: yes
    number_sections: True
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

# Load packages and data
```{r message=FALSE, warning=FALSE}
library(janitor)
library(nlme)
library(car)
library(emmeans)
library(lme4)
library(tidyverse)
library(performance)
library(DHARMa)
library(MASS)
```


```{r}
#Source maso menos plot
source("~/Documents/projects/shade_house_experiment/codes/masomenos_plot.R")
```

```{r}
data_complete_final <- 
  read.csv("~/Documents/projects/shade_house_experiment/data/data_for_biomass_leaftraits_models.csv", header = T)

```


```{r message=FALSE, warning=FALSE}


data_complete_final$treatment <- factor(data_complete_final$treatment,
					levels = c(
							   "ambientrain", 
							   "ambientrain_nutrients",
							   "ambientrain_water",
							   "ambientrain_water_nutrients"
							   )
					)

#Cleaned data
data_rate_change <- 
    data_complete_final %>% 
    
    #Order the columns  
    dplyr::select(id,spcode,family,treatment,nfixer,init_height,
                  everything()) %>% 
  	clean_names() 

data_rate_change$nfixer <- as.factor(data_rate_change$nfixer)
data_rate_change$spcode <- as.factor(data_rate_change$spcode)


levels(unique(data_rate_change$treatment))
```



# Model used 

$$response\sim treatment*fixer\ + initial\ height\ + random(1|specie)$$

```{r}
mixed_models_formula = function(response) {
  
  #'this function takes each response variable and join it to the formula
  formula = paste(response, " ~ treatment * nfixer + init_height + (1|spcode)")
  
  glmer(as.formula(formula), family = gaussian(link = "identity"), data = data_rate_change)
}


```

# Run Models

```{r}
#Takes response variable's names 
vars <- names(data_rate_change)[7:15]
```

```{r}
vars <- purrr::set_names(vars)
```

```{r message=FALSE, warning=FALSE}
mixed_models <-  vars %>%
     map(mixed_models_formula)
```

# Model assumptions for dry weights

## Total Biomasss

__No major problems detected__

```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted as:
# “the proportion of the variance explained by the grouping structure in the 
# population”
model_performance(mixed_models$totalbiom)
```



```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=20, fig.height=12}


check_model(mixed_models$totalbiom)

```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_totalbiom <- simulateResiduals(fittedModel = mixed_models$totalbiom)
plot(simulation_output_totalbiom) 
```

```{r}
testDispersion(simulation_output_totalbiom)
```
```{r}
check_heteroscedasticity(mixed_models$totalbiom)
```

## Aboveground Biomass

__No major problems detected__


```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted 
# as “the proportion of the variance explained by the grouping structure in the 
# population”
model_performance(mixed_models$above_biom)
```


```{r,fig.align='center', fig.height=8, fig.width=16, message=FALSE, warning=FALSE}
check_model(mixed_models$above_biom)
```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_above_biom <- simulateResiduals(fittedModel = mixed_models$above_biom)
plot(simulation_output_above_biom) 
```


```{r}
testDispersion(simulation_output_above_biom)
```

```{r}
check_heteroscedasticity(mixed_models$above_biom)
```


## Belowground Biomass

__NEEDS WORK__
```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted as “the proportion of the variance explained by the grouping structure in the population”
model_performance(mixed_models$below_biom)
```


```{r,fig.align='center', fig.height=8, fig.width=16, message=FALSE, warning=FALSE}


check_model(mixed_models$below_biom)

```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_below_biom <- simulateResiduals(fittedModel = mixed_models$below_biom)
plot(simulation_output_below_biom) 
```

```{r}
testDispersion(simulation_output_below_biom)
```


```{r}
check_heteroscedasticity(mixed_models$below_biom)
```

## Stem Dry weight

__No major problems detected__

```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted as “the proportion of the variance explained by the grouping structure in the population”
model_performance(mixed_models$stem_dry_weight)
```

```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
check_model(mixed_models$stem_dry_weight)
```

```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_stem_dry_weight <- simulateResiduals(fittedModel = mixed_models$stem_dry_weight)
plot(simulation_output_stem_dry_weight) 
```


```{r}
testDispersion(simulation_output_stem_dry_weight)
```


```{r}
check_heteroscedasticity(mixed_models$stem_dry_weight)
```


## Leaf Dry weight
__No major problems detected__

```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted as “the proportion of the variance explained by the grouping structure in the population”
model_performance(mixed_models$whole_leaf_dry_weight)

```


```{r,fig.align='center', fig.height=8, fig.width=16, message=FALSE, warning=FALSE}

check_model(mixed_models$whole_leaf_dry_weight)

```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_whole_leaf_dry_weight <- simulateResiduals(fittedModel =
															 	mixed_models$whole_leaf_dry_weight)
plot(simulation_output_whole_leaf_dry_weight) 
```


```{r}
testDispersion(simulation_output_whole_leaf_dry_weight)
```

```{r}
check_heteroscedasticity(mixed_models$whole_leaf_dry_weight)
```


# Model assumptions for Mass Fractions

## Root Mass Fraction 

__NEEDS WORK__

```{r}

# Model performance

# ICC provides information on the explained variance and can be interpreted as “the proportion of the variance explained by the grouping structure in the population”
model_performance(mixed_models$rmf)
```


```{r,fig.align='center', fig.height=8, fig.width=16, message=FALSE, warning=FALSE}
check_model(mixed_models$rmf)
```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_rmf <- simulateResiduals(fittedModel = mixed_models$rmf)
plot(simulation_output_rmf) 
```


```{r}
testDispersion(simulation_output_rmf)
```

```{r}
check_heteroscedasticity(mixed_models$rmf)
```


## Stem Mass fraction
__NEEDS WORK__

```{r}
# Model performance

# ICC provides information on the explained variance and can be interpreted as “the proportion of the variance explained by the grouping structure in the population”
model_performance(mixed_models$smf)
```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
check_model(mixed_models$smf)
```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 
simulation_output_smf <- simulateResiduals(fittedModel = mixed_models$smf)
plot(simulation_output_smf) 

```


```{r}
testDispersion(simulation_output_smf)
```

```{r}
check_heteroscedasticity(mixed_models$smf)
```

## Leaf Mass fraction

__No major problems detected__
```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
# Model performance

# ICC provides information on the explained variance and can be interpreted as “the proportion of the variance explained by the grouping structure in the population”
model_performance(mixed_models$lmf)

check_model(mixed_models$lmf)
```


```{r,message=FALSE, warning=FALSE, fig.align='center', fig.width=16, fig.height=8}
#check model assumptions with DHARMa 
#passes all tests so just go with it. 

simulation_output_lmf <- simulateResiduals(fittedModel = mixed_models$lmf)
plot(simulation_output_lmf) 
```


```{r}
testDispersion(simulation_output_lmf)
```

```{r}
check_heteroscedasticity(mixed_models$lmf)
```





